### 2023-03-10 Manuscript/Draft1.Rmd
updates

(see Table \@ref(tab:CodERROR)) In the Cod results, when using reduced strata adding covariates produces worst VAST results (though still better than stratified mean). VAST without covariates much worse than stratified mean in fall with increasing temperature and all strata, but adding covariates corrects this.

(see Table \@ref(tab:HadERROR)) For Haddock, VAST has a particularly hard time in spring regularly producing larger errors than the stratified mean with added covariates only improving to the level of the stratified mean. VAST shows improved results in fall relative to the stratified mean with added covariates producing extremely low errors in some cases.

In considering the Yellowtail Flounder results in Tables \@ref(tab:YTresultsallstrata) and \@ref(tab:YTresultsreducedstrata), we can see VAST estimates generally provide lower errors relative to those derived from the stratified mean, with models that include covariate information typically providing the lowest errors. However, there are several instances in which VAST failed to provide improved abundance estimates during the fall season without covariate information, producing the largest errors seen in the Table. These errors are corrected by including covariate information allowing for an improved VAST estimate that are significantly lower than their stratified mean counterparts.

@@

A more careful visual analysis of all error plots for each species reveals that VAST estimates tend to provide error estimates that are too high, while the stratified mean estimates are, on average, too low. Additionally, model/estimate ratio plots for all VAST estimates remain much closer to the desired value of 1 compared to stratified mean estimates, which can range as high as 10 and exhibit large yearly changes.

>>>>

(see Table \@ref(tab:CodERROR)) In the Cod results, when using reduced strata in constant temperature, adding covariates produces worse VAST results (though still better than stratified mean). This is similar to what we see with all of Haddock results (adding covariates makes it worse with reduced strata). This is NOT seen with Yellowtail (adding covariates always helps). VAST without covariates much worse than stratified mean in fall with increasing temperature, but adding covariates corrects this and makes VAST estimates better than stratified mean.

(see Table \@ref(tab:HadERROR)) For Haddock, VAST has a particularly hard time in spring regularly producing larger errors than the stratified mean with added covariates only improving to the level of the stratified mean. VAST shows improved results in fall relative to the stratified mean with added covariates producing extremely low errors in some cases. In Haddock results, adding covariates improves estimates only when all strata are included (when reduced strata, adding covariates actually makes VAST worse). Since this occurs in all scenarios, it seems to again be related to failing to accurately monitor strata 17.

In considering the Yellowtail Flounder results in Tables \@ref(tab:YTresultsallstrata) and \@ref(tab:YTresultsreducedstrata), we can see VAST estimates generally provide lower errors relative to those derived from the stratified mean, with models that include covariate information typically providing the lowest errors. However, there are several instances in which VAST failed to provide improved abundance estimates during the fall season without covariate information, producing the largest errors seen in the Table. These errors are corrected by including covariate information allowing for an improved VAST estimate that are significantly lower than their stratified mean counterparts. Also for Yellowtail, with all strata included adding covariates greatly improved estimates. With strata excluded there was still an improvement with covariates, but much less dramatic than with all strata.

@@

<font size="5">*Model / Estimate Ratio Results*</font>

A simple visual analysis of all error plots for each species reveals that VAST estimates tend to provide abundance estimates that are above the model value, while the stratified mean estimates are, on average, too low. This can be further examined by viewing model/estimate ratio plots, in which VAST estimates remain much closer to the desired value of 1 compared to stratified mean estimates, which can range as high as 10 and exhibit large yearly changes.

A more careful analysis of individual yearly model/estimate ratios when all strata are included revealed that about 26\% of all VAST ratios were above 1 (84\% less than 1), with an average of 85\% with a standard deviation of 0.09. On the other hand, 67\% of stratified mean estimates were above 1 with an average of 1.51 and a standard deviation of 0.31. There were seasonal differences in estimate ratios with spring VAST ratios producing (0.96,0.07), fall VAST ratios being (0.74,0.14), stratified mean spring ratios resulting in (1.31,0.19), and fall values of (1.78,0.51). The breakdown for individual species followed a similar pattern.

When the entire domain is sampled, adding covariates brings the estimate ratio closer to 1. However, with a reduced domain adding covariates can decrease the accuracy of VAST estimates. When including all strata, the stratified mean produced (1.54,0.32), while adding covariates improved the VAST estimate ratio from (0.79,0.06) to (0.92,0.06). With a reduced number of strata, the stratified mean resulted in our worst summary values of (3.48,1.82), and adding covariates to our VAST models decreased the ratio results from (1.33,0.31 ) to (1.47,0.34).

### 2023-03-10 Manuscript/Draft1.tex
updates

(see Table \ref{tab:CodERROR}) In the Cod results, when using reduced strata adding covariates produces worst VAST results (though still better than stratified mean). VAST without covariates much worse than stratified mean in fall with increasing temperature and all strata, but adding covariates corrects this.

(see Table \ref{tab:HadERROR}) For Haddock, VAST has a particularly hard time in spring regularly producing larger errors than the stratified mean with added covariates only improving to the level of the stratified mean. VAST shows improved results in fall relative to the stratified mean with added covariates producing extremely low errors in some cases.

In considering the Yellowtail Flounder results in Tables \ref{tab:YTresultsallstrata} and \ref{tab:YTresultsreducedstrata}, we can see VAST estimates generally provide lower errors relative to those derived from the stratified mean, with models that include covariate information typically providing the lowest errors. However, there are several instances in which VAST failed to provide improved abundance estimates during the fall season without covariate information, producing the largest errors seen in the Table. These errors are corrected by including covariate information allowing for an improved VAST estimate that are significantly lower than their stratified mean counterparts.

>>>>

(see Table \ref{tab:CodERROR}) In the Cod results, when using reduced strata in constant temperature, adding covariates produces worse VAST results (though still better than stratified mean). This is similar to what we see with all of Haddock results (adding covariates makes it worse with reduced strata). This is NOT seen with Yellowtail (adding covariates always helps). VAST without covariates much worse than stratified mean in fall with increasing temperature, but adding covariates corrects this and makes VAST estimates better than stratified mean.

(see Table \ref{tab:HadERROR}) For Haddock, VAST has a particularly hard time in spring regularly producing larger errors than the stratified mean with added covariates only improving to the level of the stratified mean. VAST shows improved results in fall relative to the stratified mean with added covariates producing extremely low errors in some cases. In Haddock results, adding covariates improves estimates only when all strata are included (when reduced strata, adding covariates actually makes VAST worse). Since this occurs in all scenarios, it seems to again be related to failing to accurately monitor strata 17.

In considering the Yellowtail Flounder results in Tables \ref{tab:YTresultsallstrata} and \ref{tab:YTresultsreducedstrata}, we can see VAST estimates generally provide lower errors relative to those derived from the stratified mean, with models that include covariate information typically providing the lowest errors. However, there are several instances in which VAST failed to provide improved abundance estimates during the fall season without covariate information, producing the largest errors seen in the Table. These errors are corrected by including covariate information allowing for an improved VAST estimate that are significantly lower than their stratified mean counterparts. Also for Yellowtail, with all strata included adding covariates greatly improved estimates. With strata excluded there was still an improvement with covariates, but much less dramatic than with all strata.
